\begin{abstract}

    % Riding the wave of recent groundbreaking achievements, artificial intelligence (AI) is currently the buzzword on everybody's lips and, building upon the notion of intelligence as learning, Machine Learning (ML) emerged as its pinnacle.
    % Riding the wave of recent groundbreaking achievements, artificial intelligence (AI) is currently the buzzword on everybody's lips and, addressing problems where human-crafted solutions are unfeasible, Machine Learning (ML) emerged as its pinnacle.
    % Artificial Intelligence (AI) stands as a transformative force, riding the wave of recent groundbreaking achievements, and Machine Learning (ML) emerged as its pinnacle, addressing problems where human-crafted algorithms are unfeasible and allowing algorithms to learn from historical data and discover solutions autonomously.
    % In particular, it addresses problems where human-crafted algorithms are unfeasible, allowing algorithms to learn from historical data and discover solutions autonomously.
    Riding the wave of recent groundbreaking achievements, artificial intelligence (AI) is currently the buzzword on everybody's lips and, allowing algorithms to learn from historical data, Machine Learning (ML) emerged as its pinnacle.
    The multitude of algorithms, each with unique strengths and weaknesses, highlights the absence of a universal solution and poses a challenging optimization problem.
    In response, automated machine learning (AutoML) navigates vast search spaces within minimal time constraints.
    By lowering entry barriers, AutoML emerged as promising the democratization of AI, yet facing some challenges.

    In \textbf{data-centric AI}, the discipline of systematically engineering data used to build an AI system, the challenge of configuring data pipelines is rather simple.
    We devise a methodology for building effective data pre-processing pipelines in supervised learning as well as a data-centric AutoML solution for unsupervised learning.

    In \textbf{human-centric AI}, many current AutoML tools were not built around the user but rather around algorithmic ideas, raising ethical and social bias concerns.
    % Motivated by such a mitigation call, this thesis contributes to human-centered AutoML, aiming at complementing, instead of replacing, human intelligence.
    We contribute by deploying AutoML tools aiming at complementing, instead of replacing, human intelligence.
    In particular, we provide solutions for single-objective and multi-objective optimization and showcase the challenges and potential of novel interfaces featuring large language models.

    Finally, there are application areas that rely on numerical simulators,
    % with domain experts manually configuring them through several trials and errors.
    often related to earth observations, they tend to be particularly high-impact and address important challenges such as climate change and crop life cycles.
    We commit to coupling these physical simulators with (Auto)ML solutions towards a \textbf{physics-aware AI}.
    Specifically, in precision farming, we design a smart irrigation platform that: allows real-time monitoring of soil moisture, predicts future moisture values, and estimates water demand to schedule the irrigation.

    % starting from supervised and unsupervised learning. The former is provided with a ground
    % truth to drive the learning towards a user-specified objective function

    % for both supervised learnand unsupervised learning, a
    % towards a more human-centered AutoML process.
    % In particular,
    % This places a mitigation call a shift toward the human-centered paradigm
    % Enabling human intervention, translating into the need for a shift toward the human-centered paradigm.
    % pushed towards a more human-centered AutoML process aimed at complementing, instead of replacing, human intelligence.

\end{abstract}