\begin{acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

Before diving into the depths of my research, I would like to acknowledge the sung and unsung heroes of this PhD journey.

\vspace{0.2cm}

% I would like to express my sincere gratitude to my advisors Prof. Matteo Golfarelli and Prof. Stefano Rizzi for their support and inspiration.
% My Ph.D. has been a joyful journey under their supervision.\\
The success of this thesis is certainly due to my research group, the business intelligence group (BIG).
First of all, I would like to thank my supervisor Prof. Matteo Golfarelli, who motivated me in the first place to start this PhD, and provided constant support and feedback.
I express my sincere gratitude to Prof. Stefano Rizzi for his priceless advice and encouragement.
I would like to thank my advisor Dr. Matteo Francia for his invaluable guidance, but also for welcoming me to the group ``so nicely''.
As to countless cups of coffee and \emph{italian passeggiate}, I thank my lab mates Dr. Enrico Gallinucci, Chiara Forresi, Manuele Pasini, Anna Giulia Leoni, Nicola Santolini, and Sterenn Roux.

\noindent At the Universitat Polit\`{e}cnica de Catalunya, Barcelona, huge thanks go to Prof. Alberto Abelló and Dr. Besim Bilalli for their inspiration and support during my Master's thesis, and throughout the whole PhD.

\noindent From Leibniz University, Hanover, I want to thank the AutoML group.
I would like to express my sincere gratitude to Prof. Marius Lindauer and Dr. Alexander Tornede, not only for the endless research insights and opportunities, but also for their kindness in welcoming me to the group.
Besides, I am grateful to have found an incredible group with which I had an amazing time.
I would like to thank Carolin Benjamins, Difan Deng, Theresa Eimer, Helena Graf, Aditya Mohan, Tim Ruhkopf, Sarah Segel, Daphne Theodorakopoulos, and Tanja Tornede.

\vspace{0.2cm}

With no doubts, I must thank my friends who both supported me during the ups and downs, and celebrated my achievements.
I would like to thank Gianluca Aguzzi,  Matteo Magnani, and Giuseppe Pisano.
I sincerely would not change anyone of you on this journey.
I would like to thank all the friends encountered during these years at the Cesena campus, especially in the Area 4.0 lab.
I believe the are not many groups having such a bond (and fun).
Especially, I would like to thank Giovanni Ciatto, Nicolas Farabegoli, Angela Cortecchia, Davide Domini, Matteo Magnini, Danilo Pianini, Martina Baiardi, Stefano Novellani, and Ruslan Shaiakhmetov.

\noindent My biggest thanks go to my mum Emilia for her never-ending support, solely a true \emph{mammone} can understand how much she means to me.
No less, I thank my father Renato and brother Yuri, for their continuous encouragement and for believing in me.

\noindent Last but not least, I would thank my partner Maria for cheering me up when ``it is not the day'', for her every-day caring, for all the fun, but -- mostly -- for always making me feel loved.

\end{acknowledgements}
